---
title: "20019803_A1"
output: html_notebook
---


**Load data and dplyr**
```{r}
#don't do this: 
#setwd("/Users/kelli/Desktop/Biol343/Assignment_1") 
#decDat=read.csv("DecodonAsexualityData.csv")

#do this:
decDat=read.csv("/Users/kelli/Desktop/Biol343/Assignment_1/DecodonAsexualityData.csv")

library(dplyr)
```

**Check class, dimensions, head and tail for errors**
```{r}
class(decDat)
dim(decDat)
head(decDat)
tail(decDat)

```

**Add the ovary area measure to the dataset**
```{r}

#don't do this: decDat$OvaryAreaMM <- decDat$OvaryWidthMM*decDat$OvaryHeightMM
#do this:
decDat <- decDat %>% mutate(ovary_area_mm2 = ovary_height_mm * ovary_width_mm)
```

**Calculate the minimum, maximum, mean, median and standard deviation of ovary area and flower number**
Don't do this
"""
decDat %>% filter(!is.na(FlowerNum))  %>%  summarise( min_OA = min(OvaryAreaMM), min_FN = min(FlowerNum), max_OA = max(OvaryAreaMM), max_FN = max(FlowerNum), avg_OA = mean(OvaryAreaMM), avg_FN = mean(FlowerNum), med_OA = median(OvaryAreaMM), med_FN = median(FlowerNum), sd_OA = sd(OvaryAreaMM), sd_FN = sd(FlowerNum)
)
"""
do this:
```{r}
select(decDat, ovary_area_mm2, flower_num) %>% summary()


```


**Calculate the mean ovary area, flower number, and August temperature for sexual and asexual populations**

#dont do this:
decDat %>% filter(!is.na(FlowerNum))  %>% filter(PopType=="S") %>% summarise(meanOA_s = mean(OvaryAreaMM), meanFN_s = mean(FlowerNum), meanAT_s = mean(MeanTempC))

Do this:
```{r}

?group_by
decDat %>% group_by(pop_type)  %>% summarise(meanOA = mean(ovary_area_mm2), meanFN = mean(flower_num, na.rm = T), meanAT = mean(mean_temp_c))


```
**Draw some conclusions with regards to the hypothesis you are testing.**
**Is there a potentially confounding factor that a full analysis of these data should take into account?**

Sexual populations produce 54% more flowers than asexual populations. Interestingly, the ovary area of the average asexually produced flower is  ~3% larger than that of the average sexually produced flower. Considering the small sample size (n=25), this second difference is likely not statistically significant.

It is worth considering that local variables like altitude or even latitude based circadian rythm shifts could be playing a role in dictating flowering amount and potentially reproductive type determination.